CMSAC_reveal_js

### 2025-10-10 CMSAC_reveal_js.qmd
Reveal_js Presentation

---
title: "Assessing Ranking Fidelity of Competition Structures via Weighted Mutual Information"
author: "Zach Culp, Josie Peterburs, Dr. Gregory J. Matthews, Dr. Ryan McShane"
date: "`r Sys.Date()`"
bibliography: references.bib
fontsize: 12pt
output:
  beamer_presentation:
    theme: default
---

@@

## Introduction
<!-- Overall comments 
I think the presentation is too long for 12 minutes (+ questions), but
I feel like I didn't cover everything we want. What is plausable to cover
in a 12-minute period? 
Also, we need to figure out the notation for the final metric because I think
the weight should have the subscript m, not I. And I probably was not consistent
with the usage of theta or calling it true rank/true strength. -->

@@

## New Slide - Extreme Tournament

- 4 team single elimination tournament:
  - Dream Team vs Work Group vs Middle School Girls Team vs Babies
- If we only care about the top team winning, seeding structure does not matter
- If we care about who gets second and third, seeding becomes important
  - Would not pair up Work Group and Dream Team in the first round

## New

- Comparing different structures like round robin and single eliminations, which is best?

- Efficacy: Final rankings should match the true rankings of the strength of the competitors

- Effectivity: Interested in correctly ranking the best competitor as first

CITEE

## New

- Other competitions where the correct ranking of some, but not all, of the competitors is of import
  - Olympic competition ranks the top 3 competitors

- "Salient ranks": the ranks of a competition structure that matter
  - Efficacy: 1 through $n$ matter
  - Effectivity: 1 is only salient rank

- We will propose a metric that quantifies how effectively different tournament structures convey information about competitors’ true strengths

@@

## Tournament Simulations
- Each competition structure is simulated given the number of teams, $n$,  and $\theta$
- Below is an example of the output from one simulation of a single elimination tournament structure

@@

- The notation for mutual information is: 
$$ I(X;Y) = \sum_{y \in Y}\sum_{x \in X} p(x,y) \log{\frac{p(x,y)}{p(x)p(y)}} $$
- In our case, it can be simplified to:
$$ I(r, \hat{r}) = n! \sum_{\hat{r}} p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{(\frac{1}{n!})^2}} = n! \log{((n!)^2)} $$

## Weighted Mutual Information
- A limitation of mutual information is that it does not distinguish between the direction of information
- In other words, a ranking of $(1,2,3,4)$ is equivalent to $(4,3,2,1)$
- Because of this, we proposed a weighting function $w(r, \hat{r})$
- So, the new formula is:
$$ wI_m(r,\hat{r}) = \sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}} $$

## Weighting Function
- In our original paper submission, we used squared error loss as the weighting function:
$$ w(r, \hat{r}) =
\begin{cases}
\frac{1}{\hat{r}'r}, & r \neq \hat{r} \\
1, & r = \hat{r}
\end{cases} $$
- However, after comments and reviews from people at CSAS and CMSAC, we have decided to use a weighted Kendall's Tau

@@

$$W_{ij} = w_{\min(i,j)}, \text{ for } i,j = m,2m, \dots, n\times m$$

- $m$: vector of length $n$ representing the "salient weights", or the importance of correctly getting each ranking

@@

- So, the final formula for the metric is:
$$ \frac{wI_m(r,\hat{r})}{H(r) + H(\hat{r})} = \frac{\sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}}}{\log_2((n!)^2)}$$

## Results

<!-- Should these plots be on separate slides? -->
```{r out.width="49%", fig.show='hold', fig.align='center'}
knitr::include_graphics("tournament_comparison_plot.png")
#knitr::include_graphics("tournament_comparison_plot_close.png")
```

>>>>

---
title: "CMSAC_reveal_js"
author: "Zach Culp, Josie Peterburs, Dr. Gregory J Matthews, Dr. Ryan P.A. McShane"
format: revealjs
editor: visual
---

@@

##

## Introduction

```{=html}
<!-- Overall comments 
I think the presentation is too long for 12 minutes (+ questions), but
I feel like I didn't cover everything we want. What is plausable to cover
in a 12-minute period? 
Also, we need to figure out the notation for the final metric because I think
the weight should have the subscript m, not I. And I probably was not consistent
with the usage of theta or calling it true rank/true strength. -->
```

@@

@@

## Tournament Simulations

-   Each competition structure is simulated given the number of teams, $n$, and $\theta$
-   Below is an example of the output from one simulation of a single elimination tournament structure

@@

-   The notation for mutual information is: $$ I(X;Y) = \sum_{y \in Y}\sum_{x \in X} p(x,y) \log{\frac{p(x,y)}{p(x)p(y)}} $$

<!-- -->

-   In our case, it can be simplified to:

    $$ I(r, \hat{r}) = n! \sum_{\hat{r}} p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{(\frac{1}{n!})^2}} = n! \log{((n!)^2)} $$

## Weighted Mutual Information

-   A limitation of mutual information is that it does not distinguish between the direction of information
-   In other words, a ranking of $(1,2,3,4)$ is equivalent to $(4,3,2,1)$
-   Because of this, we proposed a weighting function $w(r, \hat{r})$
-   So, the new formula is: $$ wI_m(r,\hat{r}) = \sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}} $$

## Weighting Function

-   In our original paper submission, we used squared error loss as the weighting function: $$ w(r, \hat{r}) =
    \begin{cases}
    \frac{1}{(r-\hat{r})^2}, & r \neq \hat{r} \\
    1, & r = \hat{r}
    \end{cases} $$

-   However, after comments and reviews from people at CSAS and CMSAC, we have decided to use a weighted Kendall's Tau

@@

$$W_{ij} = w_{\min(i,j)}, \text{ for } i,j = m,2m, \dots, n\times m$$
- $m$: vector of length $n$ representing the "salient weights", or the importance of correctly getting each ranking

@@

- So, the final formula for the metric is:
$$ \frac{wI_m(r,\hat{r})}{H(r) + H(\hat{r})} = \frac{\sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}}}{\log_2((n!)^2)}$$
## Results

<!-- Should these plots be on separate slides? -->
```{r out.width="49%", fig.show='hold', fig.align='center'}
knitr::include_graphics("tournament_comparison_plot.png")
knitr::include_graphics("tournament_comparison_plot_close.png")
```
